Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-09-15 Ben_Manuscripts/structure_paper/Draft.tex
revisions

%MRS5: could be punchier. I will keep thinking.  
\title{Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia}
\author{Richard D. Noble}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Joe Yelk}
\author{Matthew A. Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\begin{document}
%MRS13: we might end up adjusting the author list a bit; I will ask Matt where he thinks he and Joe should go.  I will talk to Rich when we get this going; not clear that he actually has contributed.  But that's my job. 
  \graphicspath{{./figures/}}

@@

  Nanostructured porous membranes made from the cross-linked hexagonal phase
  of self-assembled lyotropic liquid crystals (LLCs) are a promising material
  for selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular,
  we compare simulated X-ray diffraction (XRD) patterns with experimental XRD
  to quantify and understand the differences between simulation and experiment.
  We find that pores are likely composed of 5 columns of stacked LLC monomers
  which surround each hydrophilic core. We learned that these columns likely 
  move independently of each other over time scales which we can not simulate.
  We find that some of the structure previously attributed to monomer tail 
  tilt is likely instead due to ordered tail packing. Although this system
  has been reported as dry, we show that small amounts of water are necessary
  to fully reproduce all features from the experimental XRD pattern due to
  asymmetries introduced by hydrogen bonds between monomer head groups and
  water molecules. We explore the composition and structure of the nanopores
  and reveal that there exists a composition gradient rather than an abrupt
  partition between the hydrophilic and hydrophobic regions. 
  % BJC13: If this is to be included here, it needs some rewording. I allude to
  % this when talking about the independently moving columns above. 
  %A caveat is
  %that the time scales of the dynamics are extremely long for this system, 
  %resulting in simulated structures that appear too ordered which requires
  %careful examination of the metastable states observed in order to draw any
  %conclusions. 
  %MRS14: OK, let's talk through it.  I feel that this is an important caveat to add so
  %that people don't think we are claiming more than we are, and therefor should be mentioned
  % explicitly, rather than the implication left above.
  The clearer picture of the nanoscopic structure of these membranes provided
  in this study will enable a better understanding of the mechanisms of small
  molecule transport within these nanopores.

@@

  % MRS8: intro still seems a bit limited, describe more generally
  % (and briefly!) other things selective membranes are good for. The
  % paper is primarily about selective membranes, not fracking
  % water. See elimalech paper for reviews, and add a number of more
  % citations.

  %MRS14: I feel this introduction spends a bit too much time talking at too general a level.  The paper introduction to the problem tackled in the paper.  One simple way to tighten it up would be to improve the connection between the thesis and the data.  You say that highly selective membranes would be useful, but then don't provide evidence that they are the solution to the problems listed.

  % In order to perform complex separations x, y and z at reduced costs, we need highly selective membranes 
  % include all descriptions of separation processes into this paragraph

  More highly selective nanoporous membranes would be extremely useful in order
  to perform complex aqueous separations with seawater and various sources of
  wastewater. Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007}, and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} are
  examples of the diverse constituents of these water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  it is is possibl to reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus
  lowering energy requirements. Additionally, one can extract valuable resources
  from the feed streams. For example, flowback water produced during hydraulic
  fracturing of shale formations contains dissolved species such as acetate whose
  extraction has economic value \cite{dischinger_application_2017}.

@@

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, as well as ions. However the porous architecture of
  commercially 
  %MRS15: may be too strong an assertion.
  %dominant
  available 
  membranes lack uniformity which severely impacts their
  selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures which offer separation pathways that are tortuous and polydisperse
  in size \cite{werber_materials_2016}. Tortuosity increases the effective
  length that a solute must travel, while polydispersity limits membrane
  selectivity.

  % BJC14: (3) justification for why nanostructured membranes are a good idea for this
  % long-term route towards highly selective membranes are those that are nanostructured
  %MRS15: popilar for separations applications would imply the are getting a larger market share now, 
  %though my impression is that they are research topics of heavy interest, but not yet that commercially viable.  Check my chainge below.
  Nanostructured porous membrane materials have become 
  %increasingly popular
  an increasingly popular research area for
  aqueous separations applications because they offer the ability to control pore
  architecture at the molecular scale, thereby permitting the design of
  solute-specific separation membranes \cite{humplik_nanostructured_2011}.
  However their development has been limited by the ability to synthesize and
  scale 
  %MRS15: not sure what you mean by ``fundamentally sound technologies''. THat the underlying principles are good, but that scaling is hard? Fundamentally sound seems to suggest that both principles and scaling are good, but that's not what the implication is.
  various fundamentally sound technologies. Graphene sheets are atomically
  thick which results in excellent water permeability but defects during
  manufacturing severely impact selectivity \cite{cohen-tanugi_multilayer_2016}.
  Molecular dynamics (MD) simulations of carbon nanotubes show promise
  \cite{humplik_nanostructured_2011}, but synthetic techniques are unable to
  achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Consequently, there is a
  need for a scalable nanostructured porous membrane.

@@

%BJC13: You can stop here, I haven't changed anything below.

@@

%MRS8: preliminary evidence has shown they may be?  It's more than a guess.
  Preliminary evidence has shown that crosslinked lyotropic liquid crystal
  (LLC) membranes are scalable and may be capable of performing highly selective
  separations. LLCs are amphiphilic molecules that have the ability to
  self-assemble into porous nanostructures \cite{smith_ordered_1997} and can be
  crosslinked to create mechanically strong membrane films with pores on the
  order of 1 nm in diameter \cite{zhou_supported_2005}. Unlike most commercial NF
  membranes, LLC membrane pores are uniform in size because they are
  self-assembled. Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  examined in this paper are salts, and therefore lead to Donnan exclusion of
  ions in solution. The membrane gains a net surface charge when counterions from
  the head groups that line the pore walls escape into the feed solution in an
  effort to balance the gradients of concentration and electric potential
  \cite{donnan_theory_1995}.

@@

  The H\textsubscript{II} phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  c.a.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I} phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II} phase membranes represent the ideal geometry for
  achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.  However, the
  hexagonally packed liquid crystalline domains of the H\textsubscript{II} phase
  generally form isotropically oriented and therefore mutually unaligned domains
  which are detrimental to membrane permeability. This domain scale misalignment
  have inhibited further development of the this technology, and research efforts
  were focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment~\cite{dischinger_papers}.

@@

  However, recently researchers have learned how to macroscopically align the
  hexagonal domains which has revived research into H\textsubscript{II} phase LLC
  membranes. In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
  hexagonal domains, created by the monomer Na-GA3C11, using a magnetic field
  with subsequent cross-linking to lock the structure in
  place\cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could
  obtain the same result by confining neat monomer between PDMS or glass
  substrates since hexagonal mesophases preferentially anchor perpendicular to
  both surfaces\cite{feng_thin_2016}.
%  the same result using a second technique termed soft confinement\cite{feng_thin_2016}.
  %MRS8: maybe say what it is, not just what it's named? 
  Current experimental efforts are focused on extending the method to the
  H\textsubscript{II} phase and characterizing the performance of these newly
  aligned systems.

@@

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.eps}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge_thick.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{columns.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other with
	short range order and assemble into pores. (e) The hydrophilic head groups 
	(blue) facing towards the pore center. The pores assemble into hexagonally
	packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

  % BJC14: (5) justify need for simulations
  % BJC14: dishchinger example seems like a good thing to bring up since they 
  % explicitly state in the paper that they need a better understanding of
  % transport
  Our current understanding of the molecular details of LLC membranes'
  nanostructure is not sufficient to be able to precisely design them for
  specific separations. Dischinger et al. attempted to use an empirical model
  that correlates the physiochemical properties of the counterion used in a 
  Q\textsubscript{I} phase LLC membrane to solute rejection\cite{dischinger_effect_2017}.
  Although their model showed some qualitative agreement with experiment, the
  quality of fit of their model was limited due to complex solute-membrane 
  interactions which could not easily be modeled. Additionally, they observed
  an unexpected discrepancy in the relationship between uncharged solute
  rejection and water permeability which will require a more in-depth knowledge of
  the difference between solute and solvent transport.

@@

  %BJC14: nothing new after this point

@@

	\item \textit{R-$\pi$}: The location of the first is at $q_z$ = 1.7
	\AA$^{-1}$, corresponding to a real space separation of 3.7 {\AA}. Previous
	work~\cite{feng_scalable_2014} attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or z-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

@@

	\item \textit{R-pores}: The final feature corresponds to the spacing
	and symmetry of the pore columns. 
        %d\textsubscript{100} plane. This plane is geometrically
        %related the distance between pores.
	The feature, which we named R-pores, is characterized by reflections along the
	equatorial axis defined by $q_z$ = 0. The spacing between dots is indicative of
	the hexagonal symmetry of the packed pores. We observe the same information
	with higher resolution using SAXS (Fig.~\ref{fig:SAXS}).

@@

  	We are careful to avoid the term 'layers' since many liquid crystalline systems
 	have long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range z-direction correlations between 
        stacked columns of monomers. In this study, we use atomistic molecular modeling
        to study how the system's structure is affected by the density of these monomer columns
        that surround each pore's hydrophilic core.

@@

	While the Col\textsubscript{h} phase is described as dry, it is likely
	that small amounts of ambient water are absorbed into the system. The hydrogen
	bonding network formed by the water may play a role in structuring the pore. We
	used simulated X-ray diffraction (XRD) patterns to uncover any 
        %see if there is any
	meaningful structural difference between a "dry" and a "wet" system.

@@

  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system. Previous work has shown
  a coarse-grained model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see section S-\ref{S-section:self_assembly}). 
  To bypass the slow self-assembly process, we use Python scripts to assemble monomers
  into a structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:build_procedure}).
  %MRS13: Would be nice to say we put the scripts on GitHub.  Additional details about which scripts can be in supporting.
  These initial structures may result in equilibration into slowly-intercoverting metastable states, an issue we will address later in the paper. 
  % BJC5: This needs to be modified to reflect columnar build procedure
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{monomer_diagonal.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.42\textwidth}
  \includegraphics[width=\linewidth]{stacked_monomers.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.35\textwidth}
  \includegraphics[width=\linewidth]{initial_pore.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{initial_unit_cell.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) We assembled monomers into columns by stacking them on top of each 
		other. (c) We duplicated each column and rotated them to create hydrophilic pore centers.
		We chose to stack twenty monomers into each column. (d) We duplicated the pores and 
		placed them into a monoclinic unit cell with hexagonal symmetry.}\label{fig:build_procedure}
  \end{figure}

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. We find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The pore radius is chosen to be 0.5 nm in our initial
  configurations. The initial distance between pores, within a wide range, also has 
  little effect on the the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We 
  chose an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm. A sensitivity analysis of both parameters is presented in the 
  %MRS13: be consistent on capitalization of ``Supporting'' ``Information'' and ``section''. Look at JPCB for guidance.
  Supporting information, section S-\ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We also extensively explored a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this largest spacing yields non-physical behavior which is detailed in the 
  Supporting Information, section S-\ref{S-section:initial_dbwl}.

@@

  The number of columns per pore is unknown, as stated in question
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of columns per pore. We built systems in the sandwiched and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

  \subsection{Equilibration}\label{section:equilibration}
  We developed equilibration schemes for creating dry and wet configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in section S-\ref{S-section:equilibration}.

@@

>>>>

%MRS5: could be punchier. I will keep thinking.  
\title{Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes}
\author{Benjamin J. Coscia}
\author{Joe Yelk}
\author{Matthew A. Glaser}
\affiliation{Department of Physics, University of Colorado Boulder, Boulder CO, 80309, USA}
\author{Richard D. Noble}
\author{Douglas L. Gin}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\alsoaffiliation{Department of Chemistry and Biochemistry, University of Colorado Boulder, Boulder CO 80309, USA}
\author{Xunda Feng}
\affiliation{Department of Chemical and Environmental Engineering, Yale University, New Haven, Connecticut 06511, USA}
\author{Michael R. Shirts}
\email{michael.shirts@colorado.edu}
\affiliation{Department of Chemical and Biological Engineering, University of Colorado Boulder, Boulder, CO 80309, USA}
\begin{document}
%MRS13: we might end up adjusting the author list a bit; I will ask Matt where he thinks he and Joe should go.  I will talk to Rich when we get this going; not clear that he actually has contributed.  But that's my job. 
  \graphicspath{{./figures/}}

@@

  Nanostructured porous membranes made from the cross-linked hexagonal phase
  of self-assembled lyotropic liquid crystals (LLCs) are a promising material
  for selective separations. In this work, we investigate an experimentally
  characterized LLC membrane using atomistic molecular modeling. In particular,
  we compare simulated X-ray diffraction (XRD) patterns with experimental XRD
  to quantify and understand the differences between simulation and experiment.
  We find that pores are likely composed of 5 columns of stacked LLC monomers
  which surround each hydrophilic core. We learned that these columns likely 
  move independently of each other over time scales which we can not simulate.
  We find that some of the structure previously attributed to monomer tail 
  tilt is likely instead due to ordered tail packing. Although this system
  has been reported as dry, we show that small amounts of water are necessary
  to fully reproduce all features from the experimental XRD pattern due to
  asymmetries introduced by hydrogen bonds between monomer head groups and
  water molecules. We explore the composition and structure of the nanopores
  and reveal that there exists a composition gradient rather than an abrupt
  partition between the hydrophilic and hydrophobic regions. A caveat is
  that the time scales of the dynamics are extremely long for this system, 
  resulting in simulated structures that appear too ordered which requires
  careful examination of the metastable states observed in order to draw any
  conclusions. The clearer picture of the nanoscopic structure of these 
  membranes provided in this study will enable a better understanding of
  the mechanisms of small molecule transport within these nanopores.

@@

  More highly selective nanoporous membranes would be extremely useful in order
  to perform complex aqueous separations with seawater and various sources of
  wastewater. Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007}, and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} are
  examples of the diverse constituents of these water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  it is possible to reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus
  lowering energy requirements. Additionally, one can extract valuable resources
  from the feed streams. For example, flowback water produced during hydraulic
  fracturing of shale formations contains dissolved species such as acetate whose
  extraction has economic value \cite{dischinger_application_2017}.

@@

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of 1 nm
  in size and smaller, as well as ions. However the porous architecture of
  commercially available membranes lack uniformity which severely impacts their
  selectivity \cite{van_der_bruggen_review_2003}. Although scalable, their
  fabrication involves the spontaneous assembly of polymers into disordered
  structures which offer separation pathways that are tortuous and polydisperse
  in size \cite{werber_materials_2016}. Tortuosity increases the effective
  length that a solute must travel, while polydispersity limits membrane
  selectivity.

  % BJC14: (3) justification for why nanostructured membranes are a good idea for this
  % long-term route towards highly selective membranes are those that are nanostructured

  Nanostructured porous membrane materials have become an increasingly popular
  research area for aqueous separations applications because they offer the 
  ability to control pore architecture at the molecular scale, thereby 
  permitting the design of solute-specific separation membranes 
  \cite{humplik_nanostructured_2011}. However their development has been 
  limited by the ability to synthesize and scale various technologies. Graphene
  sheets are atomically thick which results in excellent water permeability but
  defects during manufacturing severely impact selectivity \cite{cohen-tanugi_multilayer_2016}.
  Molecular dynamics (MD) simulations of carbon nanotubes show promise
  \cite{humplik_nanostructured_2011}, but synthetic techniques are unable to
  achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Consequently, there is a
  need for a scalable nanostructured porous membrane.

@@

@@

  Preliminary evidence has shown that crosslinked lyotropic liquid crystal
  (LLC) membranes are scalable and may be capable of performing highly selective
  separations. LLCs are amphiphilic molecules that have the ability to
  self-assemble into porous nanostructures \cite{smith_ordered_1997} and can be
  crosslinked to create mechanically strong membrane films with pores on the
  order of 1 nm in diameter \cite{zhou_supported_2005}. Unlike most commercial NF
  membranes, LLC membrane pores are uniform in size because they are
  self-assembled. Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  examined in this paper are salts, and therefore lead to Donnan exclusion of
  ions in solution. The membrane gains a net surface charge when counterions from
  the head groups that line the pore walls escape into the feed solution in an
  effort to balance the gradients of concentration and electric potential
  \cite{donnan_theory_1995}.

@@

  The H\textsubscript{II} phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  The H\textsubscript{II} phase forms at room temperature in the presence of
  c.a.~10 wt\% water and consists of hexagonally packed, hydrophilic pore
  columns\cite{smith_ordered_1997}. In the absence of water, neat monomer will
  form the same hexagonal columnar structure which, in literature, has been
  referred to as the Col\textsubscript{h} thermotropic
  phase\cite{feng_scalable_2014}. 
  % BJC6: maybe not a necessary detail
%  The most
%  promising Q\textsubscript{I} phase for membrane applications forms at 70\degree
%  C when monomer is mixed with c.a.~20 wt\% glycerol\cite{carter_glycerol-based_2012}.
  Q\textsubscript{I} phase membranes consist of a tortuous network of three
  dimensionally interconnected pores that prevent optimal through-plane
  transport. In contrast, the densely packed, non-tortuous and uniform sized
  pores of H\textsubscript{II} phase membranes represent the ideal geometry for
  achieving high solute flux\cite{matyka_tortuosity-porosity_2008}.  However, the
  hexagonally packed liquid crystalline domains of the H\textsubscript{II} phase
  generally form isotropically oriented and therefore mutually unaligned domains
  which are detrimental to membrane permeability. This domain scale misalignment
  had inhibited further development of this technology, and research efforts
  were focused on the Q\textsubscript{I} phase, whose geometry does not require
  alignment~\cite{zhou_new_2007}.

@@

  However, recently researchers have learned how to macroscopically align the
  hexagonal domains which has revived research into H\textsubscript{II} phase LLC
  membranes. In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
  hexagonal domains, created by the monomer Na-GA3C11, using a magnetic field
  with subsequent cross-linking to lock the structure in
  place\cite{feng_scalable_2014}. In 2016, Feng et al. showed that one could
  obtain the same result by confining neat monomer between PDMS or glass
  substrates since hexagonal mesophases preferentially anchor perpendicular to
  both surfaces\cite{feng_thin_2016}. Current experimental efforts are focused
  on extending the method to the H\textsubscript{II} phase and characterizing
  the performance of these newly aligned systems.

@@

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.eps}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.png}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=\textwidth]{wedge_thick.png}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{columns.png}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=\textwidth]{hexagonal_packing.png}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other to
	create columns with short range order then assemble into pores with  
	hydrophilic head groups (blue) face towards the pore center. (e) The 
	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

  % BJC14: (5) justify need for simulations

  Our current understanding of the molecular details of LLC membranes'
  nanostructure is not sufficient to be able to precisely design them for
  specific separations. Dischinger et al. attempted to use an empirical model
  that correlates the physiochemical properties of the counterion used in a 
  Q\textsubscript{I} phase LLC membrane to solute rejection\cite{dischinger_effect_2017}.
  Although their model showed some qualitative agreement with experiment, the
  quality of fit of their model was limited due to complex solute-membrane 
  interactions which could not easily be modeled. Additionally, they observed
  an unexpected discrepancy in the relationship between uncharged solute
  rejection and water permeability which will require a more in-depth knowledge of
  the difference between solute and solvent transport.

@@

@@

	\item \textit{R-$\pi$}: The location of the first is at $q_z$ = 1.7
	\AA$^{-1}$, corresponding to a real space separation of 3.7 {\AA}. Previous
	work attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or z-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

@@

	\item \textit{R-pores}: The final feature corresponds to the spacing
	and symmetry of the pore columns. The feature, which we named R-pores, 
	is characterized by reflections along the equatorial axis defined by $q_z$ = 0.
	The spacing between dots is indicative of the hexagonal symmetry of the 
	packed pores, and the location of the leading peak (closest to $q_r$ = 0)
	is related to the distance between pores. We observe the same information
	with higher resolution by looking at the same system's 1D SAXS pattern 
	(Fig.~\ref{fig:SAXS}).

@@

  	We are careful to avoid the term ``layers" since many liquid crystalline systems
 	have long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range z-direction correlations between 
        stacked columns of monomers. In this study, we use atomistic molecular modeling
        to study how the system's structure is affected by the density of these monomer columns
        that surround each pore's hydrophilic core.

@@

	While the Col\textsubscript{h} phase is described as dry, it is likely
	that small amounts of ambient water are absorbed into the system. The hydrogen
	bonding network formed by the water may play a role in structuring the pore. We
	used simulated X-ray diffraction (XRD) patterns to uncover any 
        %see if there is any
	meaningful structural difference between a ``dry" and a ``wet" system.

@@

  The timescale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  calling for a more efficient way to build the system. Previous work has shown
  a united-atom model self assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see section S-\ref{S-section:self_assembly}). 
  To bypass the slow self-assembly process, we use Python scripts to assemble monomers
  into a structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:build_procedure}).
  %MRS13: Would be nice to say we put the scripts on GitHub.  Additional details about which scripts can be in supporting.
  These initial structures may result in equilibration into slowly-intercoverting
  metastable states, an issue we will address later in the paper.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{monomer_diagonal.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.42\textwidth}
  \includegraphics[width=\linewidth]{stacked_monomers.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.35\textwidth}
  \includegraphics[width=\linewidth]{initial_pore.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{initial_unit_cell.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) We assembled monomers into columns by stacking them on top of each 
		other. (c) We duplicated each column and rotated them to create hydrophilic pore centers.
		We chose to stack twenty monomers into each column. (d) We duplicated the pores and 
		placed them into a monoclinic unit cell with hexagonal symmetry.}\label{fig:build_procedure}
  \end{figure}

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. The equilibrium
  configuration is sensitive to some while insensitive to others. We find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The pore radius is chosen to be 0.5 nm in our initial
  configurations. The initial distance between pores, within a wide range, also has 
  little effect on the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We 
  chose an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental value
  of 4.12 nm. A sensitivity analysis of both parameters is presented in the 
  %MRS13: be consistent on capitalization of ``Supporting'' ``Information'' and ``section''. Look at JPCB for guidance.
  Supporting information, section S-\ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with monomers placed so aromatic rings stack 3.7 \AA~apart in
  the $z$-direction. We also extensively explored a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this largest spacing yields non-physical behavior which is detailed in the 
  Supporting Information, section S-\ref{S-section:initial_dbwl}.

@@

  The number of columns per pore is unknown, as stated in Question
  (\ref{point:monomernum}). We tested configurations constructed with a varied
  number of columns per pore. We built systems in the sandwiched and parallel
  displaced configurations with 4, 5, 6, 7 and 8 monomers per layer.

  \subsection{Equilibration}\label{section:equilibration}

  We developed equilibration schemes for creating dry and wet configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in section S-\ref{S-section:equilibration}.

@@

  \subsubsection*{Shape and Intensity of the $q_r$ cross-section of R-$\pi$}
